---
title: "Chemical Space"
author: "Chuleeporn Phanus-umporn, Saw Simeon, Watshara Shoombuatong and Chanin Nantasenamat"
date: "July 13, 2559 BE"
output: pdf_document
---

## Chemical space of anti-sickling
```{r, warning=FALSE, message=FALSE, error=FALSE, cache = TRUE}
library(cowplot)
data <- read.csv("Linpiski_Descriptors.csv")
descriptors_name <- c("ALogP", "nHBAcc", "nHBDon", "MW")
descriptors <- data[, descriptors_name]
Activity <- data$Activity
data_2 <- cbind(Activity, descriptors)
data_plot <- data.frame(data_2)
p <- ggplot(data_plot, aes(MW, ALogP))
p <- p + geom_point(aes(colour = factor(Activity)), size = 3, alpha = 0.7)
p <- p + theme(legend.position = ("none"),
                panel.border = element_rect(linetype = "solid", 
                                            colour = "black", fill = NA, size = 1),
                axis.text.x = element_text(colour = "black", size = 10),
                axis.text.y = element_text(colour = "black", size = 10),
                plot.margin = grid::unit(c(1, 1, 1, 1), "cm"),
                axis.title.x = element_text(colour = "black", size = 15, face = "bold"),
                axis.title.y = element_text(colour = "black", size = 15, face = "bold"))
print(p)

```

\pagebreak

## Visualization of Lipinski's descriptors

### Chemical space of AChE inhibitors are shown as active (green), inactive (red) and intermediate (blue)

```{r, warning= FALSE, message=FALSE, cache = TRUE}
library(ggplot2)
library(cowplot)
data_2 <- cbind(Activity, descriptors)
data_2 <- na.omit(data_2)
data_plot <- data.frame(data_2)
p <- ggplot(data_plot, aes(MW, ALogP))
p <- p + geom_point(aes(colour = factor(Activity)), size = 3, alpha = 0.3)
p <- p + theme(legend.position = ("none"),
           panel.border = element_rect(linetype = "solid",
                                       colour = "black", fill = NA, size = 1),
           axis.text.x = element_text(colour = "black", size = 10),
           axis.text.y = element_text(colour = "black", size = 10),
           plot.margin = unit(c(1, 1, 1, 1), "cm"),
           axis.title.x = element_text(colour = "black", size = 15, face = "bold"),
           axis.title.y = element_text(colour = "black", size = 15, face = "bold")
           )
print(p)
```

\pagebreak

### Boxplot of Lipinski's rule-of-five descriptors (ALogP) 

```{r, warning= FALSE, message=FALSE, cache = TRUE}

p_1 <- ggplot(data_plot, aes(factor(Activity), ALogP))
p_1 <- p_1 + geom_boxplot(aes(fill = factor(Activity)), alpha = 0.7)
p_1 <- p_1 + theme(legend.position = ("none"),
                   panel.border = element_rect(linetype = "solid",
                                               colour = "black", fill = NA, size = 1),
                   axis.text.x = element_text(colour = "black", size = 15),
                   axis.text.y = element_text(colour = "black", size = 15),
                   plot.margin = unit(c(1, 1, 1, 1), "cm"),
                   axis.title.y = element_text(size = 20, face = "bold"),
                   axis.title.x = element_blank())

print(p_1)
```

\pagebreak

### Boxplot of Lipinski's rule-of-five descriptors (nHBAcc) 

```{r, warning= FALSE, message=FALSE, cache = TRUE}

p_2 <- ggplot(data_plot, aes(factor(Activity), nHBAcc))
p_2 <- p_2 + geom_boxplot(aes(fill = factor(Activity)), alpha = 0.7)
p_2 <- p_2 + theme(legend.position = ("none"),
                   panel.border = element_rect(linetype = "solid",
                                               colour = "black", fill = NA, size = 1),
                   axis.text.x = element_text(colour = "black", size = 15),
                   axis.text.y = element_text(colour = "black", size = 15),
                   plot.margin = unit(c(1, 1, 1, 1), "cm"),
                   axis.title.y = element_text(size = 20, face = "bold"),
                   axis.title.x = element_blank())
print(p_2)
```

\pagebreak

### Boxplot of Lipinski's rule-of-five descriptors (nHBDon) 

```{r, warning= FALSE, message=FALSE, cache = TRUE}

p_3 <- ggplot(data_plot, aes(factor(Activity), nHBDon))
p_3 <- p_3 + geom_boxplot(aes(fill = factor(Activity)), alpha = 0.7)
p_3 <- p_3 + theme(legend.position = ("none"),
                   panel.border = element_rect(linetype = "solid",
                                               colour = "black", fill = NA, size = 1),
                   axis.text.x = element_text(colour = "black", size = 15),
                   axis.text.y = element_text(colour = "black", size = 15),
                   plot.margin = unit(c(1, 1, 1, 1), "cm"),
                   axis.title.y = element_text(size = 20, face = "bold"),
                   axis.title.x = element_blank())
print(p_3)
```

\pagebreak

### Boxplot of Lipinski's rule-of-five descriptors (MW) 

```{r, warning= FALSE, message=FALSE, cache = TRUE}


p_4 <- ggplot(data_plot, aes(factor(Activity), MW))
p_4 <- p_4 + geom_boxplot(aes(fill = factor(Activity)), alpha = 0.7)
p_4 <- p_4 + theme(legend.position = ("none"),
                   panel.border = element_rect(linetype = "solid",
                                               colour = "black", fill = NA, size = 1),
                   axis.text.x = element_text(colour = "black", size = 15),
                   axis.text.y = element_text(colour = "black", size = 15),
                   plot.margin = unit(c(1, 1, 1, 0.4), "cm"),
                   axis.title.y = element_text(size = 20, face = "bold"),
                   axis.title.x = element_blank()) 

print(p_4)
```

\pagebreak

# Arranging boxplot of Lipinski's rule-of-file descriptors (ALogP, nHBAcc, nHBDon, MW)
```{r, fig.width = 10, fig.height= 10, warning=FALSE, message=FALSE, cache = TRUE}
library(cowplot)
plot_grid(p_1, p_2, p_3, p_4)

```



## EDA of feature important analysis

```{r, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE, fig.align = 'center', cache = TRUE}

library(caret)
library(reshape2)

df <- read.csv("data.csv")
names <- c("Activity", "SubFPC171", "SubFPC5", "SubFPC1", "SubFPC298", "SubFPC2", "SubFPC16", "SubFPC173", "SubFPC18", "SubFPC275", 
           "SubFPC100")
df <- df[, names]
data_melt <- melt(df)
ggplot(aes(y = value, x = variable, fill = Activity), data = data_melt) + geom_boxplot() + 
  ylab("Substructure Count") + theme(
    legend.position = "none",
    axis.text.y = element_text(size = 20, colour = "black"),
    axis.text.x = element_text(size = 20, colour = "black"),
    plot.margin = unit(c(1,1, 1, 1),"cm"),
    panel.border = element_rect(linetype = "solid", colour = "black", fill = NA, size = 1),
    axis.title = element_text(size = 25, face = "bold", colour = "black")
  )



```

